\textbf{ Matching the forecast horizon with the relevant ecological processes }

### 2019-07-30 doc/manuscript.tex
Edited Intro

A recent flurry of synthesis papers in high impact journals attests to surging interest in ecological forecasting. Many of the papers 
argue that a commitment to ecological forecasting is necessary to guide the environmental policy and management
decisions that will become increasingly critical under global change \citep{clark_ecological_2001,mouquet_review:_2015,dietze_iterative_2018}.
A second argument is that rigorous testing of predictions will accelerate the process of basic research and discovery 
\citep{houlahan_priority_2017,dietze_prediction_2017,dietze_iterative_2018}.
Taken together, these papers, along with a new textbook \citep{dietze_ecological_2017} represent an emerging consensus that 
forecasting should be a priority for ecological research.

No such consensus is available for exactly how to tackle this challenge.
Ecologists rely on many different approaches and models for making predictions, and 
frequently argue about the virtues of process-based vs. phenomenological models (refs?).
We often justify investments in basic research by claiming that improved process-level understanding will improve predictions (e.g.,\citealt{Adler2012}),
and it is easy to find examples of how extrapolation of linear relationships leads to poor predictions under novel conditions \citep{Fitzpatrick,Veloz2012}.
But in many cases simple phenomenological models make better predictions than complex process-based models with high parameter
uncertainty (refs? \citealt{tredennick_we_2017}?).

The pragmatic solution to these debates is clear: we should forecast using the models that make the best predictions.
Identifying these models requires rigorously testing predictions against observations. However, many common ecological predictions,
such as a species' geographic range in the year 2100, are virtually impossible to test (but see \citealt{Maguire2015}), 
leading \citep{dietze_iterative_2018} to call for greater emphasis on near-term forecasting. The assumption is that building 
our capacity to make short-term forecasts will ultimately improve our long-term forecasts.

Our goal is to consider a hypothesis that, in our view, has not been carefully considered by either the push for near-term forecasting or
ongoing debates about competing modeling approaches. We expect that short and long-term forecasts will require models
incorporating different kinds of processes and fit to different kinds of data sets. We demonstrate this idea using two simulation case studies, the first 
focused on how competitive interactions affect population dynamics of a focal species, the second for  an 
eco-evolutionary scenario. Our analyses show that: 
\begin{enumerate}
	\item For short-term forecasts, phenomenological time-series approaches are hard to beat, whereas longer-term forecasts require understanding slow 
	processes such as evolutionary and ecological selection as well as dispersal.
	\item Different kinds of data reflect the operation of different processes: longitudinal data capture autocorrelation and responses of current assemblages, while 
	data spanning spatial gradients capture the long-term outcome of selection and dispersal.  Whether predictive models should be trained using longitudinal or
	spatial data sets, or both, depends on the time-scale of the desired forecast.
	\item A key challenge for future research is determining the rate at which slow processes first influence and ultimately dominate dynamics.
\end{enumerate}

>>>>

Forecasting is increasingly recognized as important to the application and advancement of ecological research.
Forecasts are necessary to guide environmental policy and management
decisions about mitigation and adaption to global change \citep{clark_ecological_2001,mouquet_review:_2015,dietze_iterative_2018}.
But forecasts can also advance understanding of the processes governing ecological systems by providing rigorous tests of 
model predictions \citep{houlahan_priority_2017,dietze_prediction_2017,dietze_iterative_2018}.
The combined benefits of informing policy and management and advancing basic research 
makes forecasting an important priority for ecological research.

The models used for ecological forecasting have traditionally been based on either time-series approaches or space-for-time substitutions.
The time-series approach involves fitting models to long-term datasets to describe the temporal dynamics of a system.  We then use those
dynamic models to make predictions about what will happen in the future. This is the common approach for population forecasts and includes 
both process-based and data-driven models (e.g., \citep{ward_complexity_2014}). Whether process-based or
data-driven, time-series models typically capture processes operating on interannual time-scales, such as birth, death, 
small scale dispersal events, and short-term responses to environmental conditions.
Space-for-time substitution approaches begin by describing how an ecological variable of interest, such as occupancy or productivity, 
varies across sites experiencing different environmental conditions. These spatial relationships between environment and ecological 
response are assumed to also hold for changes at a site through time. To make a forecast, we first predict the future
environmental conditions and then determinie the associated ecological response, based on the observed spatial relationship. 
Most forecasts for future species distributions and biodiversity are based on this approach. 
Space-for-time models capture processes operating over long time periods, such as
immigration, extinction, and responses to large or prolonged environmental changes. Although both of these approaches are
widely used, there has been little discussion of their advantages and disadvantages for guiding policy decisions or 
advancing our understanding of ecological dynamics \citep{harris_forecasting_2018} (but see Renwick et al ??).

Whether historical dynamics, contemporary spatial patterns, or spme combination of the two will serve as the best source of 
information for forecasting may depend on how far into the future we are attempting to forecast \citep{harris_forecasting_2018}. 
This potential dependency on the "forecast horizon"; ()\textit{sensu} citation)
reflects lags in the response of ecological conditions to environmental change, shifts in the importance of ecological processes with
time scale \citep{levin_1992,rosenzweig_1995}, and differences between time-series and spatial gradients in the range of 
environmental conditions represented in observed data.
At short forecast horizons (days to years), dynamics will reflect the physiological and demographic responses of 
the organisms present at a site more than than turnover of genotypes or species, 
environmental conditions are likely to stay within the range of historical variation, 
and the current state of the system is likely to capture the influence of unmeasured processes. As a result, 
for near-term forecasts time-series approaches may capture the key dynamics and provide accurate predictions.

In contrast, at long forecast horizons (decades to centuries),  environmental conditions that have not been 
historically observed are likely to not only occur but to persist long enough to drive significant turnover of genotypes and species 
along with changes in the flux of energy and nutrients.  At these long scales, the current state of the system may be 
little help in predicting the future state. For the century-scale forecasts often featured in biodiversity and 
species-distribution modeling, space-for-time approaches may effectively capture the response of ecosystems to major shifts 
in climate over long periods, producing better long-term forecasts than time-series approaches. 
Using different modeling approaches for different forecast horizons is common in other disciplines.
For example, meteorological models for short-term weather forecasts differ substantially in spatial
and temporal resolution and extent from the global circulation models used to predict long-term changes
in climate [IS THIS REALLY TRUE? REFERENCE?].

Here we use simulation models to 1) demonstrate that the best model-building approaches for ecological forecasting 
depend on the time horizon of the forecast, and 2) explore how time-series and space-for-time approaches might be 
combined to make better forecasts at intermediate time scales. We conduct two simulation case studies, one 
focused on how interspecific interactions affect the population dynamics of a focal species, and the second focused on
an eco-evolutionary scenario. Our analyses show that: 
\begin{enumerate}
	\item For short-term forecasts, phenomenological time-series approaches are hard to beat, whereas longer-term forecasts require understanding slow processes such as evolutionary and ecological selection as well as dispersal.
	\item Different kinds of data reflect the operation of different processes: longitudinal data capture autocorrelation and responses of current assemblages, while data spanning spatial gradients capture the long-term outcome of selection and dispersal.  Whether predictive models should be trained using longitudinal or spatial data sets, or both, depends on the time-scale of the desired forecast.
	\item A key challenge for future research is determining the rate at which slow processes first influence and ultimately dominate dynamics.
\end{enumerate}
